---
title: "Project 2"
author: "Skyler Moomey"
date: "`r Sys.Date()`"
output: 
  html_document:
    df_print: paged
    fig_caption: true
    highlights: pygments
    number_sections: no
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

The goal of this project is to apply Simple Linear Regression (SLR) to a dataset on cameras. The variables of interest in this study are maximum camera resolution and lowest camera resolution setting. By applying the theoretical underpinnings of SLR to the data, verifying required assumptions about the SLR model, and calculating the extent to which the model fits the data, we verify that there is indeed a positive linear correlation between our variables.

# Introduction

## The Data

```{r}
camera_dt <- read.csv("camera_dataset.csv");
head(camera_dt[,3:4])
```


The data I am using come from a Kaggle dataset on 1038 different camera models, which include the variables max resolution and low-resolution settings. These variables describe the maximum resolution in pixels of a photo that a camera is capable of taking and the lowest resolution setting of the camera. My statistical hypothesis is that the maximum resolution will increase without a comparable increase in the lowest resolution of the camera software, and I will be using a SLR to determine whether there actually is a statistical relationship between the variables.

## Data Cleaning

```{r}
length(camera_dt)
length(camera_dt$Max.resolution)
low <- camera_dt$Low.resolution

camera_dt <- camera_dt[low > 0,]

length(camera_dt$Max.resolution)

##removed all values at low=0 due to several of them being outliers
```

Note: A measure of 0 pixels as a lowest resolution does not have value as a measure. If some camera has a 0 pixel lowest resolution, then no photo would be rendered. For this reason, I am removing all experimental units with a lowest resolution on file of 0 pixels, as this is probably just a stand-in value for an uncertain lowest resolution rating. Therefore, of the 1038 experimental units, only 984 will be used in my study.

## Preliminary Plots and Interpretation of the Data

```{r}
library (s20x)
pairs20x(camera_dt[,3:4])

library(ggplot2)
g = ggplot(camera_dt, aes(y=Max.resolution, x= Low.resolution)) + geom_point()
g = g + geom_smooth(method = "loess")
g
```

Contrary to my hypothesis, it does appear that there is a linear relationship between lowest resolution setting and maximum camera resolution. The data appear to appear to have a linear correlation.

# Theoretical Basis of SLR

A Simple Linear Regression (SLR) is a predictive, probabilistic model for a sample. Since we believe a linear relationship exists between Lowest Camera Resolution setting and Maximum Camera Resolution, we can create an equation of a line to estimate values of Maximum Camera Resolution, our dependent variable. Letting $y$ represent maximum resolution and $x$ represent lowest resolution setting, we create the following equation:

Consider $y_i$, the value of a dependent variable in the sample.

We can describe $y_i$ as:

$$
\begin{equation}
  \begin{split}
  y_i &= \beta_0 + \beta_1 x_i + \epsilon_i
  \end{split}
\end{equation}
$$

In this view, $\hat y = \beta_0 + \beta_1 x_i$ represents our predictive model, a linear equation, and $\epsilon_i$ is the error term in our model. When we add the error term to $\hat y$, we recover the true value of our sample datum, $y_i = \hat y_i + \epsilon_i$. Obviously, $x_i$ will be the dependent variable. $\beta_0$ and $\beta_1$ are unknown constants that we will find values of to minimize our error terms. We know that some amount of error must be present in the model, but according to assumption, $E(\epsilon) = 0$. This validly applies, as a line running through the center of the data will have error in both the positive and negative directions. If the line is a good fit, the mean value of these error terms will be zero. As such,

$$
\begin{equation}
  \begin{split}
  E(y) = \bar y &= E(\beta_0 + \beta_1 x_i + \epsilon_i)\\
  & = E(\beta_0) + E(\beta_1 x_i) + E(\epsilon_i)\\
  & = \beta_0 + \beta_1 \bar x + 0 \\
  & = \beta_0 + \beta_1 \bar x
  \end{split}
\end{equation}
$$
Thus, we can see that $\bar y$ is in terms of $\bar x$, an unbiased estimator, so we use this equation as a point estimate of $y$ at each datum.

## Equation

```{r}
lin.mod <- with(camera_dt, lm(Max.resolution~Low.resolution))

summary(lin.mod)$coef
```

Finally, in creating our linear model, we use the method of least squares built in to the R function lm(). This creates an equation such that error terms are minimized based on its SSR (sum of squares of residuals). The function estimates them to a $95 \%$ confidence. we find $\hat \beta_0 = 720.1612$ and $\hat \beta_1 = 0.9490$. Our p-values for both estimators are infinitesimally small, so we can assume that they are validly estimated. Our final equation is

$$
\begin{equation}
y_i = 720.1612 + 0.9490 x_i
\end{equation}
$$

# Assumptions

For an SLR to be a valid linear model, four assumptions about the data must be met:

## Linear Relationship

The first assumption of an SLR is that there is a linear relationship between the independent and dependent variables. As we can see based on our loess plot, there appears to be a roughly linear relationship between Low.resolution and Max.resolution. To assure ourselves of this fact, we apply a straight line to the data.

```{r}
with(camera_dt, plot(Max.resolution~Low.resolution, bg="Red",pch=23, cex=1.2, ylim=c(0,1.2*max(Max.resolution)), xlim=c(0,1.2*max(Max.resolution)), main="Scatter Plot and Linear Model of Max v. Low Resolution"))

lin.mod <- with(camera_dt, lm(Max.resolution~Low.resolution))
abline(lin.mod)
```

As can be seen from applying a straight to the data, it does appear that a linear relationship is present between the variables, though we will perform more analyses to determine the model's predictive accuracy.

## Linear Independence of Residuals

The second assumption for SLR is that the residuals, that is, the error terms in a linear model, are independent. For this to be the case, there should not be a steady increase or decrease in the length of the residuals over the course of the data, which would imply that early residuals are affecting later residuals.

```{r}
with(camera_dt, plot(Max.resolution~Low.resolution,bg="black", pch=21,cex=1.2))

maxhat = with(camera_dt, predict(lin.mod, data.frame(Low.resolution)))
with(camera_dt, segments(Low.resolution, Max.resolution, Low.resolution, maxhat, col="red"))

abline(lin.mod)
```

As we can see from our graph (residuals in red), there does not seem to be a constant increase or decrease in the length of our residuals. For this reason, we can assume that they are independently distributed.

## Constant Variance (Homoscedasticity)

The third assumption that must be met for a linear model to be an accurate predictor of a population is that the residuals have a constant variance throughout the sample and be distributed around a mean of zero. This can be determined by making a graph that matches our fitted $\hat y$ estimates to the true values in the sample.

```{r}
plot(lin.mod, which=1)
```
As we can see from the Residuals v. Fitted plot, our errors are likely normally distributed since they seem to have an almost constant slope through the x-axis. Our linear model is supported as a model of best-fit for the data.

## Normality

The final assumption that must be met for a SLR model to be valid is normality of the distribution of the residuals. This can also be verified using the previous plot, as a constant distribution about 0 implies that our residuals are normally distributed. To visualize this better, we can also examine a Q-Q plot of the residuals (Shapiro-Wilk Normality Test).

```{r}
normcheck(lin.mod, shapiro.wilk = TRUE)
title("Normal Distribution")
```

# Analysis of Data

A model's $R^2$ is the ratio of RSS (residual sum of squares) to TSS (total sum of squares). The closer an $R^2$ value to 1, the better the model fits the data. The third value of concern is MSS (mean sum of squares), due to the equation $TSS = MSS + RSS$. As such, for $MSS/TSS=1$, $RSS$ to be minimized as much as possible.

## RSS

The Residual Sum of Squares can be visualized with the previous figure, that of the data points and their distance from the regression line. We represent the figure here, for convenience.

```{r}
res <- residuals(lin.mod)


with(camera_dt, plot(Max.resolution~Low.resolution,bg="black", pch=21,cex=1.2, main="Residual Sum of Squares Visualization"))

maxhat = with(camera_dt, predict(lin.mod, data.frame(Low.resolution)))
with(camera_dt, segments(Low.resolution, Max.resolution, Low.resolution, maxhat, col="red"))

abline(lin.mod)
```

The formula for calculating RSS is given by $RSS=\sum_{1}^{n}(y_i - \hat y_i)^2 = \sum_{1}^{n}\epsilon_i^2$. This turns out to be,

```{r}
RSS = with(camera_dt, sum((maxhat - Max.resolution)^2))
RSS
```

## MSS

MSS (mean sum of squares) can be visualized as the distance of the mean of the sample $y_i's$ from the mean of the population, estimated using the linear model. It can be visualized with the following graph.

```{r}
with(camera_dt, plot(Max.resolution~Low.resolution,bg="blue", pch=21,cex=1.2, main="Mean Sum of Squares Visualization"))


with(camera_dt, segments(Low.resolution, mean(Max.resolution), Low.resolution, maxhat, col="light green"))

abline(lin.mod)
abline(h=mean(camera_dt$Max.resolution))
```

The value of MSS is defined to be $MSS=\sum_{1}^{n}(\bar y - \hat y)^2$. We calculate MSS for our model as,

```{r}
MSS = with(camera_dt, sum((maxhat - mean(Max.resolution))^2))
MSS
```


## TSS

Finally, TSS (total sum of squares), can be visualized as the distance from each $y_i$ from $\bar y$. Graphically, it can be shown as,

```{r}

with(camera_dt, plot(Max.resolution~Low.resolution,bg="blue", pch=21,cex=1.2, main="Total Sum of Squares Visualization"))

with(camera_dt, segments(Low.resolution, Max.resolution, Low.resolution, mean(Max.resolution), col="green"))

abline(h=mean(camera_dt$Max.resolution))
```

The formula for TSS is $MSS = \sum_{1}^{n}(y_i - \bar y)^2$. The value we calculate for TSS is,

```{r}
TSS = with(camera_dt, sum((Max.resolution - mean(Max.resolution))^2))
TSS
```

Finally, we have everything we need to calculate $R^2$.

```{r}
R2=MSS/TSS
R2
```

Our $R^2$ being this close to $1$ shows that our model is a very close fit to the data.


# Conclusion

Given the very probable linear relationship between maximum photo resolution and lowest resolution setting, we can assume that in camera manufacturing, the maximum resolution is positively, and linearly, correlated with the lowest resolution setting. This manufacturing choice seems odd, however, it seems probable that a company creating a high resolution camera would want to maintain this high resolution by scaling its lowest resolution setting up accordingly.

# References

https://www.kaggle.com/crawford/1000-cameras-data/ 
Website from which sample data were drawn.